Compositional equivalences based on Open pNets

### 2020-04-17 2016-Weak-FH-Bisim/WeakBisim.tex
quelques TODO en plus
en pages 19 et 84, sur les \gamma de longueur >1

\paragraph{Previous Works and Contribution}
While most of our previous works relied on closed, fully-instantiated semantics~\cite{BBCHM:article2008,AmeurBoulifa2017,HKM-FASE16}, it is only recently that we could design a first version of a  parameterised semantics for pNets with a strong bisimulation equivalence~\cite{henrio:Forte2016}. This article builds open this previous parameterised semantics and provide a cleaner version of the semantics with a slightly simplified formalism and extends it to deal with weak bisimulation. The slight simplification of the formula allowed us to adopt a much cleaner presentation where the theory for open automata, the algebra used to define the semantics of pNets with holes, can be studied independently from its application to pNets. Also, in previous work the study of compositionability was only partial, and in particular the proof that bisimulation is an equivalence is one new contribution of this article and provides a particularly interesting insight on the semantic model we use.
The new formalism allowed us to extend the work and define weak bisimulation  for open automata, which is entirely new. This allows us to define a weak bisimulation equivalence for open pNets with valuable properties of compositionality. 
To summarise, the contribution of this paper are the following
\begin{itemize}
\item The definition of open automata: an algebra of parameterised automata with hole with a strong bisimulation equivalence. This is an adaptation of~\cite{henrio:Forte2016} with an additional property stating that strong bisimulation equivalence is indeed an equivalence relation.
\item A semantics for open pNets expressed as translation to open automata. This is an adaptation of~\cite{henrio:Forte2016} with a complete proof that strong bisimulation is compositional.
\item A theory of weak bisimulation for open automata, and its properties. It relies on the definition of weak open transitions that are derived from transitions of the open automaton by concatenating invisible action transitions with one (visible or not) transition. The precise definition of the concatenation is also a major contribution of this article.
\item A resulting weak bisimulation equivalence for open pNets and a simple static condition on synchronisation vectors inside pNets that is sufficient to ensure that weak bisimulation is compositional.
\item An illustrative example \TODO{TODO}
\end{itemize}

@@

As required actions are now sequences of observable actions, we need an operator to build them from set of actions that occur in open transitions, i.e. an operator that takes a set of actions performed by one hole and produces a sequence of observable actions.
Thus we define $\vis{\set\beta}$ as the mapping $\set\beta$  with only observable actions of the holes in $I$, but where each element is either empty or a list of length 1:
 \[\vis{\beta_i^{i\in I}} = [\beta_i]^{i\in I'}\text{ where }I'=\left\{i| i\in I \land \beta_i\neq \tau\right\}\]

As an example the $\vis{\set\beta}$ built from the transition $OT_1$ in page \pageref{OT:SimpleProt} is $\texttt{P}\mapsto [\texttt{p-send(m)}]$. Remark that in our simple example no $\tau$ transition involves any visible action from a hole, so we have no $\beta$ sequences of length longer than 1 in the weak automaton.

@@

>>>>

\paragraph{Previous Works and Contribution}
While most of our previous works relied on closed, fully-instantiated semantics~\cite{BBCHM:article2008,AmeurBoulifa2017,HKM-FASE16}, it is only recently that we could design a first version of a  parameterised semantics for pNets with a strong bisimulation equivalence~\cite{henrio:Forte2016}. This article builds open this previous parameterised semantics and provide a cleaner version of the semantics with a slightly simplified formalism and extends it to deal with weak bisimulation. The slight simplification of the formula allowed us to adopt a much cleaner presentation where the theory for open automata, the algebra used to define the semantics of pNets with holes, can be studied independently from its application to pNets. Also, in previous work the study of compositionability was only partial, and in particular the proof that bisimulation is an equivalence is one new contribution of this article and provides a particularly interesting insight on the semantic model we use.
The new formalism allowed us to extend the work and define weak bisimulation  for open automata, which is entirely new. This allows us to define a weak bisimulation equivalence for open pNets with valuable properties of compositionality. 
To summarise, the contribution of this paper are the following
\begin{itemize}
\item The definition of open automata: an algebra of parameterised automata with hole with a strong bisimulation equivalence. This is an adaptation of~\cite{henrio:Forte2016} with an additional property stating that strong bisimulation equivalence is indeed an equivalence relation.
\item A semantics for open pNets expressed as translation to open automata. This is an adaptation of~\cite{henrio:Forte2016} with a complete proof that strong bisimulation is compositional.
\item A theory of weak bisimulation for open automata, and its properties. It relies on the definition of weak open transitions that are derived from transitions of the open automaton by concatenating invisible action transitions with one (visible or not) transition. The precise definition of the concatenation is also a major contribution of this article.
\item A resulting weak bisimulation equivalence for open pNets and a simple static condition on synchronisation vectors inside pNets that is sufficient to ensure that weak bisimulation is compositional.
\item An illustrative example based on a simple transport protocol, showing the construction of the weak open transitions, and the proof of weak bisimulation.
\end{itemize}

@@

As required actions are now sequences of observable actions, we need an operator to build them from set of actions that occur in open transitions, i.e. an operator that takes a set of actions performed by one hole and produces a sequence of observable actions.
\TODO{Ca serai tpas mal de trouver une notation plus intuitive au lieu de ce $\vis{\set\beta}$ ?}
Thus we define $\vis{\set\beta}$ as the mapping $\set\beta$  with only observable actions of the holes in $I$, but where each element is either empty or a list of length 1:
 \[\vis{\beta_i^{i\in I}} = [\beta_i]^{i\in I'}\text{ where }I'=\left\{i| i\in I \land \beta_i\neq \tau\right\}\]

As an example the $\vis{\set\beta}$ built from the transition $OT_1$ in page \pageref{OT:SimpleProt} is $\texttt{P}\mapsto [\texttt{p-send(m)}]$. Remark that in our simple example no $\tau$ transition involves any visible action from a hole, so we have no $\beta$ sequences of length longer than 1 in the weak automaton.
\TODO{en fait c'est faux ! les boucles SS\_1, SS\_2, WS\_1, WS\_2 sont aussi valides avec p\_a ou q\_b = tau, donc on peut avoir des sequences de longueur arbitraire d'actions de P et de Q}

@@

\TODO{ajouter les boucles de $WI_1$ et $WI_2$, dans le cas ou $p_a$ (resp $q_b$) sont des $\tau$:}

$WI_{12*}(n) = \openrule
{\{(\texttt{P}\mapsto \tau|\texttt{Q}\mapsto \tau)^n\},true, ()}
  {s \OTWeakarrow {\tau} s} 
s \in S$
